---
title: "Interpret_QGComp_110223_WGCNAResults"
author: "Samantha Lapehn"
date: "2023-11-03"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Interpret WGCNA Bootstrap QGComp Results from 110123
This code will be used to interpret and visualize the WGCNA bootstrapped QGComp results. This analysis used the CANDLE WGCNA modules and the CANDLE OPE data for the 5 OPEs with >70% above LOD. 

### Load Packages
```{r packages}
library(tidyverse)
library(qgcomp)
library(data.table)
library(edgeR)

source("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/EWAS_QGComp Code from Dennis/EWAS_QGComp_Functions_Update3.R")
```

### Load QGComp BS Results object for WGCNA
```{r load BS}
load(file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/EWAS_QGComp/BoostrapResults_WGCNA_110123.RData")
```

### Pull Bootstrap Results
```{r pull results BS}
Results_BS <- ewas_qgcomp_fit.boot$results 

Results_BS$FDR<-p.adjust(Results_BS$pval, method="BH", n=length(Results_BS$pval))
Results_BS$Bonferroni <- p.adjust(Results_BS$pval, method="bonferroni", n=length(Results_BS$pval))

Results_BS_Sig <- Results_BS %>%
  dplyr::filter(pval<0.05)

#Significant Modules (p<0.05)
print(Results_BS_Sig$probeID)

```

### Pull Component Effects
```{r comp effects BS}
comp_effects_BS <- ewas_qgcomp_fit.boot$component_effects

#Filter for significant modules

comp_effects_BS_sig <- comp_effects_BS %>%
  dplyr::filter(probeID %in% c("MEgreenyellow"))

comp_effects_BS_greenyellow <- comp_effects_BS %>%
  dplyr::filter(probeID == "MEgreenyellow")
comp_effects_BS_greenyellow$OPE <- c("DEP", "DBUP & DIBP", "BCETP", "BDCPP", "DPHP")
```

### Plot Sig Module Component Effect Weights
The sum of all individual component effect sizes will sum to the overall mixture effect.
```{r plot CE}
ggplot(comp_effects_BS_greenyellow, aes(x=OPE, y=weight, fill=OPE)) + 
  geom_bar(stat="identity") + 
  theme_bw() + 
  scale_y_continuous(limits = c(-0.005, 0.005)) + 
  geom_hline(yintercept = 0) + 
  coord_flip() + 
  ggtitle("Component Weights-GreenYellow (N=737)")

# Sum overall mixture effects
sum(comp_effects_BS_greenyellow$weight)

```
