---
title: "QGComp_Results_May2024"
author: "Samantha Lapehn"
date: "2024-05-27"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## QGComp Results Evaluation - May2024

Here I am uploading the results to evaluate genes with marginal significance (p<0.001) since none of the results passed the threshold of FDR<0.1. 


### Load Packages
```{r load packages}
require(tidyverse)                  
require(qgcomp)                    
require(data.table) 
library(pheatmap)
library(RColorBrewer)

source("/active/paquette_a/QGComp_sif/EWAS_QGComp Code from Dennis/EWAS_QGComp_Functions_Update3.R")   
```

### Load Results
```{r load Results}
FullResults <-read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/OPEQGComp_Evan_2024/BootstrapResults_FullDataset_FDR_crctd_03202024.csv")

MaleResults <-read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/OPEQGComp_Evan_2024/BootstrapResults_MaleStratified_FDR_crctd_03202024.csv")

FemaleResults <-read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/OPEQGComp_Evan_2024/BootstrapResults_FemaleStratified_FDR_crctd_03192024.csv")

#Loading annotation file too
load(file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/FilteredNormalizedData_102023.RData")
```

### Filter results
Filtering for genes with p<0.001 since none met the FDR<0.1 threshold
```{r filter results}
Sig_FullResults <- FullResults %>%
  dplyr::filter(pval<0.001)

Sig_MaleResults <- MaleResults %>%
  dplyr::filter(pval<0.001)

Sig_FemaleResults <- FemaleResults %>%
  dplyr::filter(pval<0.001)

## Add annotation of gene name to Sig Results

Sig_FullResults <- inner_join(Sig_FullResults, annot, by=c("probeID"="ENSEMBL"))

Sig_MaleResults <- inner_join(Sig_MaleResults, annot, by=c("probeID"="ENSEMBL"))
```


### Load Comparison Data
Loading the following: 
1) IndivGene_DEGs- this csv file contains the genes that were associated with individual OPE compounds with FDR<0.1 for both binary and continuous OPE analyses
2) Loading a dataframe of WGCNA modules that were significantly associated with individual or mixture OPEs
3) Loading full WGCNA results for gene/color annotation
```{r Comparison Data}
IndivGene_DEGs <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/Gene Results/IndivGenesResultsforComparison_May2024.csv")

WGCNA_SigModules <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/WGCNA/WGCNAResults_forComparison_May2024.csv")

load(file="/active/paquette_a/CANDLE_WGCNA/CANDLE_wide_WGCNA_ensembl.RData")

```

### Evaluate overlap with DEG Comparison Data
No overlapping genes between QGComp (p<0.001) genes and individual OPE DEGs (FDR<0.1)
```{r evaluate overlap}
DEGs_vs_FullQGComp <- inner_join(Sig_FullResults, IndivGene_DEGs, by=c("probeID"="Ensembl.ID"))

DEGs_vs_MaleQGComp <- inner_join(Sig_MaleResults, IndivGene_DEGs, by=c("probeID"="Ensembl.ID"))
```

### Evaluate overlap with WGCNA Comparison Data
Evaluating if the genes associated with mixture OPEs (p<0.001) from the full and male models overlap with any of the WGCNA modules that were associated with individual or mixture OPEs. 
```{r WGCNA comparison}
#Pull names of module colors that were significantly associated with individual or mixture OPEs
SigModuleColors <- WGCNA_SigModules$Module
#Combine with gene annotation for those modules
WGCNAgenes <-rownames(geneModuleMembership)
gene_membership <- cbind(WGCNAgenes, moduleColors)
gene_membershipDF <- data.frame(gene_membership)
SigModuleGenes <- gene_membershipDF %>%
  dplyr::filter(moduleColors %in% SigModuleColors)
#Run overlap comparison with QGComp sig genes (p<0.001)
WGCNA_vs_FullQGComp <- inner_join(Sig_FullResults, SigModuleGenes, by=c("probeID"="WGCNAgenes"))

WGCNA_vs_MaleQGComp <- inner_join(Sig_MaleResults, SigModuleGenes, by=c("probeID"="WGCNAgenes"))
```


### Load full QGComp results objects
Loading results for the full and male analyses (not female since no p<0.001 genes)
```{r load full results object}
load(file="/active/paquette_a/QGComp_sif/CodeforInterpretation_and_FDRAdjustment/MaleStratified/Male_BootstrapResults_03192024.RData")
MaleResultsObject <- ewas_qgcomp_fit.boot

FullResultsObject <- load(file="/active/paquette_a/QGComp_sif/CodeforInterpretation_and_FDRAdjustment/QGComp_Output_Files/BootstrapResults_FullDataset_03202024.RData")
FullResultsObject <- ewas_qgcomp_fit.boot
```

### Pull Component Effects 
Need to collapse nested lists into dataframes first then pull the CE object
```{r pull CE}
MaleCollapse <-tibble(MaleResultsObject)  %>% 
      unnest_wider(MaleResultsObject)
FullCollapse <-tibble(FullResultsObject)  %>% 
      unnest_wider(FullResultsObject)
#Make Component Effects Data Frame
MaleCollapse_CE <- MaleCollapse$component_effects
MaleCE <-as.data.frame(do.call(rbind, lapply(MaleCollapse_CE, as.data.frame)))

FullCollapse_CE <- FullCollapse$component_effects
FullCE <-as.data.frame(do.call(rbind, lapply(FullCollapse_CE, as.data.frame)))

```

### Evaluate CE for Sig Genes
```{r Evaluate CE}
#Full DAta Analysis
SigFullGenes <- Sig_FullResults$probeID
Sig_FullCE <- FullCE %>%
  dplyr::filter(probeID %in% SigFullGenes)
Full_sumCE <- aggregate(weight~probeID,Sig_FullCE,sum)

#Male Data Analysis
SigMaleGenes <- Sig_MaleResults$probeID
Sig_MaleCE <- MaleCE %>%
  dplyr::filter(probeID %in% SigMaleGenes)
Male_sumCE <- aggregate(weight~probeID,Sig_MaleCE,sum)
```

### Join Module Membership to CE
```{r module membership join}
Full_Sig_CE_ModMem <- full_join(WGCNA_vs_FullQGComp, Full_sumCE, by="probeID")
Full_Sig_CE_ModMem$moduleColors <- paste0(Full_Sig_CE_ModMem$moduleColors, "_All")
Male_Sig_CE_ModMem <- full_join(WGCNA_vs_MaleQGComp, Male_sumCE, by="probeID")
Male_Sig_CE_ModMem$moduleColors <- paste0(Male_Sig_CE_ModMem$moduleColors, "_Male")
```

### Add column specifying model type
```{r}
Malevec <-replicate(36, "Male")
Fullvec <- replicate(6, "All")
Full_Sig_CE_ModMem$Model <-Fullvec
Male_Sig_CE_ModMem$Model <-Malevec
Full_Sig_CE_ModMem <- Full_Sig_CE_ModMem %>%
  dplyr::select(-X)
```

### Combine all sig results into one dataframe
```{r combine male and full}
Male_Full_Sig <- rbind(Full_Sig_CE_ModMem, Male_Sig_CE_ModMem)
```

### Write results CSV to rearrange for figure
```{r write results}
#write.csv(Male_Full_Sig, "/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/OPEQGComp_Evan_2024/Male_Full_SigGenesSummary_052924.csv")
```

### Reload the results data rearranged for the figure
```{r reload data rearranged}
Heatmapinput <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/OPEQGComp_Evan_2024/QGComp_IndivGene_HeatmapInput_053024.csv")
rownames(Heatmapinput) <- Heatmapinput$X
Heatmapinput <- Heatmapinput %>%
  dplyr::select(-X)
```

### Prepare additional heatmap input
```{r heatmap input}
colors <- rev(brewer.pal(11, "RdBu"))

# Annotation
Sex <- c("All", "Male", "All", "Male","Male", "Male", "Male", "Male", "Male", "Male", "Male", "Male")

Module <-c("grey", "grey", "greenyellow", "greenyellow", "blue", "brown", "cyan", "darkred", "plum1", "royalblue", "turquoise", "yellow")

Annotation<- data.frame("Sex"=Sex, "Module"=Module)
rownames(Annotation)<-rownames(Heatmapinput)

annotation_colors = list(
  Sex=c(All="Black",Female="deeppink",Male="Blue"),
  Module=c(grey="gray41",  greenyellow="greenyellow",  blue="#0000FF", brown="brown", cyan="cyan", darkred="darkred", plum1="plum1", royalblue="royalblue", turquoise="turquoise", yellow="yellow"))

```

### Heatmap
```{r heatmap}
Breaks<-seq(-0.5,0.5, length.out=12)
pheatmap(Heatmapinput,
         color = colors,
        breaks=Breaks,
         scale='none',# unscaled data
         na.color="grey",
         cluster_rows =F,
         cluster_cols=F,
         na_col="grey",
         cellwidth=10, #Change tehe for size
         cellheight=20, #change these for siz
         fontsize_col =12, 
         fontsize_row=0.01,
         annotation_row=Annotation, 
         annotation_colors = annotation_colors)
```

### Hub Gene Analysis
### Upload hub gene files of interest
NOt uploading grey since that is the unassigned module
```{r hub gene analysis}
greenyellow_hub <-read.csv("/active/paquette_a/CANDLE_WGCNA/module_hubGenes/greenyellow_hubGenes.csv", header=FALSE)
blue_hub <-read.csv("/active/paquette_a/CANDLE_WGCNA/module_hubGenes/blue_hubGenes.csv", header=FALSE)
brown_hub <-read.csv("/active/paquette_a/CANDLE_WGCNA/module_hubGenes/brown_hubGenes.csv", header=FALSE)
cyan_hub <-read.csv("/active/paquette_a/CANDLE_WGCNA/module_hubGenes/cyan_hubGenes.csv", header=FALSE)
darkred_hub <-read.csv("/active/paquette_a/CANDLE_WGCNA/module_hubGenes/darkred_hubGenes.csv", header=FALSE)
plum1_hub <-read.csv("/active/paquette_a/CANDLE_WGCNA/module_hubGenes/plum1_hubGenes.csv", header=FALSE)
royalblue_hub <-read.csv("/active/paquette_a/CANDLE_WGCNA/module_hubGenes/royalblue_hubGenes.csv", header=FALSE)
turquoise_hub <-read.csv("/active/paquette_a/CANDLE_WGCNA/module_hubGenes/turquoise_hubGenes.csv", header=FALSE)
yellow_hub <-read.csv("/active/paquette_a/CANDLE_WGCNA/module_hubGenes/yellow_hubGenes.csv", header=FALSE)
```

### Check for overlap with module hubs
```{r hub overlap}
greenyellow_overlap <- inner_join(greenyellow_hub, Male_Full_Sig, by=c("V1"="SYMBOL"))

blue_overlap <- inner_join(blue_hub, Male_Full_Sig, by=c("V1"="SYMBOL"))

brown_overlap <- inner_join(brown_hub, Male_Full_Sig, by=c("V1"="SYMBOL"))

cyan_overlap <- inner_join(cyan_hub, Male_Full_Sig, by=c("V1"="SYMBOL"))

darkred_overlap <- inner_join(darkred_hub, Male_Full_Sig, by=c("V1"="SYMBOL"))

plum1_overlap <- inner_join(plum1_hub, Male_Full_Sig, by=c("V1"="SYMBOL"))

royalblue_overlap <- inner_join(royalblue_hub, Male_Full_Sig, by=c("V1"="SYMBOL"))

turquoise_overlap <- inner_join(turquoise_hub, Male_Full_Sig, by=c("V1"="SYMBOL"))

yellow_overlap <- inner_join(yellow_hub, Male_Full_Sig, by=c("V1"="SYMBOL"))
```


