---
title: "QGComp"
author: "Samantha Lapehn"
date: "2023-08-17"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## QGComp for OPE Data
Using this code to run quantile g-based computation for OPE mixtures and the placental transcriptome using code shared by Dennis Khodsevich from UC-Berkeley. https://github.com/D-Khodasevich/EWAS_QGComp

### Load Packages and Functions
```{r packages}
library(tidyverse)
library(qgcomp)
library(data.table)
library(edgeR)

source("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/EWAS_QGComp Code from Dennis/EWAS_QGComp_Functions_Update3.R")

```

### Load Data
Need covariates and exposure data, as well as WGCNA data.
```{r load data}
#contains MEs
load(file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CANDLE_WGCNA_copy_110123/CANDLE_wide_WGCNA_symbol.RData")
#Contains covariate and exposure data
load(file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/FilteredNormalizedData_102023.RData")

#Select only ME data for which there is OPE exposure data
MEs_forOPE <- MEs %>%
  dplyr::filter(row.names(MEs) %in% row.names(OPE_Data_Select_Ready))

```

### Prepare Exposure Data
Need to remove binary exposures so that only the continuous OPEs are included
```{r prep exposure data}
Mixture <- OPE_Data_Select_Ready %>%
  dplyr::select(log_dep, log_dbup_dibp, log_bcetp, log_bdcpp, log_dphp)
Mixture$pathways_id <-rownames(Mixture)

Xnm <- c("log_dep", "log_dbup_dibp", "log_bcetp", "log_bdcpp", "log_dphp")
```

### Prepare Covariate Data
Need to remove any covariates that we aren't using
```{r covar prep}
Covariates_Selected <- Covariates %>%
  dplyr::select(pathways_id, h_c_sex, opfr_sg, h_m_delivery_age, h_m_prepreg_bmi, h_m_race, h_m_enroll_educ, batches, cg_labor_type, h_del_method, h_m_alcohol_prenatal, Smoking_Composite, h_firstborn, h_gravidity)

covars <- c("h_c_sex", "opfr_sg", "h_m_delivery_age", "h_m_prepreg_bmi", "h_m_race", "h_m_enroll_educ", "batches", "cg_labor_type", "h_del_method", "h_m_alcohol_prenatal", "Smoking_Composite", "h_firstborn", "h_gravidity")
```

### Combine Covariate and Exposure Data into Single dataframe
```{r combine exposure and covar}
Pheno <- inner_join(Covariates_Selected, Mixture, by=c("pathways_id"))
rownames(Pheno) <- Pheno$pathways_id
Pheno <- Pheno %>%
  dplyr::select(-pathways_id)
class(Pheno$h_c_sex)
#Make sure certain covariates are indicated as factors
Pheno$h_c_sex <- as.factor(Pheno$h_c_sex)
Pheno$h_m_race <- as.factor(Pheno$h_m_race)
Pheno$h_m_enroll_educ <- as.factor(Pheno$h_m_enroll_educ)
Pheno$batches <- as.factor(Pheno$batches)
Pheno$cg_labor_type <- as.factor(Pheno$cg_labor_type)
Pheno$h_del_method<- as.factor(Pheno$h_del_method)
Pheno$h_m_alcohol_prenatal <- as.factor(Pheno$h_m_alcohol_prenatal)
Pheno$Smoking_Composite <- as.factor(Pheno$Smoking_Composite)
Pheno$h_firstborn <- as.factor(Pheno$h_firstborn)
Pheno$h_gravidity <- as.factor(Pheno$h_gravidity)

#Check data types
summary(Pheno)

```

### Prep Expression Data
```{r expr prep}
MEs_transpose <- t(MEs_forOPE)
meth <- MEs_transpose
```

### Check that order of Expression and Pheno Data are the same
```{r check order}
check <- colnames(meth) == rownames(Pheno)
sum(check)

reorder <- match(rownames(Pheno), colnames(meth))
meth_reorder <- meth[,reorder]

check2 <- colnames(meth_reorder) == rownames(Pheno)
sum(check2)
```

## Bootstrap EWAS QGComp
will perform bootstrapping for all WGCNA modules

### Run Bootstrap EWAS QGComp
Two new parametes need to be set for bootstrap version: 

Bval is the number of bootstraps to perform
seedval is to set a seed for reproducibility

```{r bootstrap}
ewas_qgcomp_fit.boot <- ewas_qgcomp.boot(pheno = Pheno, meth = meth_reorder, 
                           mix_comp = Xnm, covars = covars, mval_conversion=FALSE, 
            qval=4, output_type="full", Bval=1000, seedval=1234)
```

### Save BS Results object
```{r save BS}
#save(ewas_qgcomp_fit.boot, file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/EWAS_QGComp/BoostrapResults_WGCNA_110123.RData")

```

### Pull Bootstrap Results
```{r pull results BS}
Results_BS <- ewas_qgcomp_fit.boot$results 

Results_BS$FDR<-p.adjust(Results_BS$pval, method="BH", n=length(Results_BS$pval))
Results_BS$Bonferroni <- p.adjust(Results_BS$pval, method="bonferroni", n=length(Results_BS$pval))

Results_BS_Sig <- Results_BS %>%
  dplyr::filter(pval<0.05)

#Significant Modules (p<0.05)
print(Results_BS_Sig$probeID)

```

### Pull Component Effects
```{r comp effects BS}
comp_effects_BS <- ewas_qgcomp_fit.boot$component_effects

#Filter for significant modules

comp_effects_BS_sig <- comp_effects_BS %>%
  dplyr::filter(probeID %in% c("MEgreenyellow"))

comp_effects_BS_greenyellow <- comp_effects_BS %>%
  dplyr::filter(probeID == "MEgreenyellow")
comp_effects_BS_greenyellow$OPE <- c("DEP", "DBUP & DIBP", "BCETP", "BDCPP", "DPHP")
```

### Plot Sig Module Component Effect Weights
The sum of all individual component effect sizes will sum to the overall mixture effect.
```{r plot CE}
ggplot(comp_effects_BS_greenyellow, aes(x=OPE, y=weight, fill=OPE)) + 
  geom_bar(stat="identity") + 
  theme_bw() + 
  scale_y_continuous(limits = c(-0.005, 0.005)) + 
  geom_hline(yintercept = 0) + 
  coord_flip() + 
  ggtitle("Component Weights-GreenYellow (N=737)")

# Sum overall mixture effects
sum(comp_effects_BS_greenyellow$weight)

```

