---
title: "RNAseq_Processing_Filtering_SexStratified"
author: "Samantha Lapehn"
date: "2023-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## RNAseq Processing and Filtering for OPE Analysis- Sex Stratified

This code performs processing and filtering of the CANDLE placental transcriptome for OPE project. Performing this code after processing the sex stratified covariates/exposure data (most recently completed on 10/23/23) so that we can filter to only include RNAseq data has full covariates and exposures data.

# Load Packages
```{r load packages}
library(tidyverse)
library(edgeR)
```

# Load Data
Load RNAseq Data and Exposure/Covariate data which may be needed for model construction

The OPE_Data_Select object contains only the relevant columns that we need. Continuous data for those that pass LOD threshold and binary data for those that do not pass LOD threshold. The other dataframe (OPE_Data) contains continuous and binary data regardless of LOD threshold .
```{r Load Data}
load(file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CANDLE_RNAseq/candle_20200226.Rdata")
Covariates_Male <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CleanCovariates_Male_102323.csv")
OPE_Data_Male <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CleanOPEData_Male_102323.csv")

OPE_Data_Select_Male <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CleanOPEData_BinaryLogContinuous_Male_102323.csv")

Covariates_Female <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CleanCovariates_Female_102323.csv")
OPE_Data_Female <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CleanOPEData_Female_102323.csv")

OPE_Data_Select_Female <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CleanOPEData_BinaryLogContinuous_Female_102323.csv")
```

### Pull Count Data
Using length scaled TPM values
```{r count data}
LS_Counts <- data.frame(candle_lengthScaledTPM$counts)
LS_Counts_Male <- LS_Counts %>%
  dplyr::select(any_of(Covariates_Male$pathways_id))

LS_Counts_Female <- LS_Counts %>%
  dplyr::select(any_of(Covariates_Female$pathways_id))
  
```

### Filtering
- 1. Remove participants that don't have full covariate data - already done in code chunk above when separating RNAseq based on Sex
-2. Removed Enzembl IDs with the suffix "_PAR_Y"  
- 3. Remove non-protein coding genes
- 4. Performed Filtering on genes with mean cpm<0  
- 5. Removed Duplicated Genes  
- 6. TMM Normalization 

```{r filtering}
#Make seperate annot for Male and Female
annot_Male <- annot
annot_Female <- annot

#Original Dimensions
dim(LS_Counts_Male)
dim(annot_Male)

dim(LS_Counts_Female)
dim(annot_Female)

#Remove Ensembl IDs with the suffix "_PAR_Y"
LS_Counts_Male<-LS_Counts_Male[!grepl("_PAR_Y",rownames(LS_Counts_Male)),]
annot_Male$ENSEMBL<-as.character(annot_Male$ENSEMBL)
annot_Male<-annot_Male[!grepl("_PAR_Y",annot_Male$ENSEMBL),]

LS_Counts_Female<-LS_Counts_Female[!grepl("_PAR_Y",rownames(LS_Counts_Female)),]
annot_Female$ENSEMBL<-as.character(annot_Female$ENSEMBL)
annot_Female<-annot_Female[!grepl("_PAR_Y",annot_Female$ENSEMBL),]


#Dimsenions after "_PAR_Y" removal
dim(LS_Counts_Male)
dim(annot_Male)

dim(LS_Counts_Female)
dim(annot_Female)

#get rownames of count dataset to match annot$ENSEMBL by removing the "." and 
# everything after it
rownames(LS_Counts_Male)<-gsub("\\..*","",rownames(LS_Counts_Male))
rownames(LS_Counts_Female)<-gsub("\\..*","",rownames(LS_Counts_Female))

#Remove non-protein coding genes
BiotypesToKeep<-c("protein_coding")

GenesToKeep_Male<-NA
for(i in 1:length(BiotypesToKeep)){
  GenesToKeep_Male=c(GenesToKeep_Male,annot_Male[which(annot_Male$BIOTYPE==BiotypesToKeep[i]),"ENSEMBL"])
}
annot_Male<-annot_Male[which(annot_Male$ENSEMBL%in%GenesToKeep_Male),]
table(annot_Male$BIOTYPE)

LS_Counts_Male<-LS_Counts_Male[which(rownames(LS_Counts_Male)%in%GenesToKeep_Male),]

GenesToKeep_Female<-NA
for(i in 1:length(BiotypesToKeep)){
  GenesToKeep_Female=c(GenesToKeep_Female,annot_Female[which(annot_Female$BIOTYPE==BiotypesToKeep[i]),"ENSEMBL"])
}
annot_Female<-annot_Female[which(annot_Female$ENSEMBL%in%GenesToKeep_Female),]
table(annot_Female$BIOTYPE)

LS_Counts_Female<-LS_Counts_Female[which(rownames(LS_Counts_Female)%in%GenesToKeep_Female),]

#Dimensions after reducing to protein-coding genes
dim(LS_Counts_Male)
dim(annot_Male)

dim(LS_Counts_Female)
dim(annot_Female)

sum(rownames(LS_Counts_Male)==annot_Male)
sum(rownames(LS_Counts_Female)==annot_Female)
#Remove genes with low expression
#Keep genes with average log cpm>0
logcpm_Male<-cpm(LS_Counts_Male,log=T)

keep_Male<-rowMeans(logcpm_Male)>0
summary(keep_Male)

logcpm_filt_Male <- logcpm_Male[keep_Male,]
LS_Counts_filt_Male<-LS_Counts_Male[keep_Male,]
annot_Male<-annot_Male[keep_Male,]

logcpm_Female<-cpm(LS_Counts_Female,log=T)

keep_Female<-rowMeans(logcpm_Female)>0
summary(keep_Female)

logcpm_filt_Female <- logcpm_Female[keep_Female,]
LS_Counts_filt_Female<-LS_Counts_Female[keep_Female,]
annot_Female<-annot_Female[keep_Female,]

#Plot before and after filtering out low expressing genes

plot(density(logcpm_Male),main="Male-Before Removing low expressing genes")#,xlim=c(-10,20000))
abline(v =0, col = "red", lty = 2)

plot(density(logcpm_filt_Male),main="Male-After Removing low expressing genes")#,xlim=c(-10,20000))
abline(v =0, col = "red", lty = 2)

plot(density(logcpm_Female),main="Female-Before Removing low expressing genes")#,xlim=c(-10,20000))
abline(v =0, col = "red", lty = 2)

plot(density(logcpm_filt_Female),main="Female-After Removing low expressing genes")#,xlim=c(-10,20000))
abline(v =0, col = "red", lty = 2)

#Dimensions after low expresion filtering
dim(LS_Counts_filt_Male)
dim(annot_Male)
dim(logcpm_filt_Male)

dim(LS_Counts_filt_Female)
dim(annot_Female)
dim(logcpm_filt_Female)

sum(rownames(LS_Counts_filt_Male)==annot_Male)
sum(rownames(LS_Counts_filt_Female)==annot_Female)

#Remove duplicated genes
summary(duplicated(logcpm_filt_Male))
summary(duplicated(logcpm_filt_Female))

dupgenes_Male<-as.character(annot_Male[annot_Male$ENSEMBL%in%rownames(logcpm_filt_Male)[duplicated(logcpm_filt_Male)],"ENSEMBL"])
dupgenes_Male
dupgenes_Female<-as.character(annot_Female[annot_Female$ENSEMBL%in%rownames(logcpm_filt_Female)[duplicated(logcpm_filt_Female)],"ENSEMBL"])
dupgenes_Female

annot_Male<-annot_Male[-which(annot_Male$ENSEMBL%in%dupgenes_Male),]
logcpm_filt_Male<-logcpm_filt_Male[-which(duplicated(logcpm_filt_Male)),]
LS_Counts_filt_Male<-LS_Counts_filt_Male[-which(duplicated(LS_Counts_filt_Male)),]

annot_Female<-annot_Female[-which(annot_Female$ENSEMBL%in%dupgenes_Female),]
logcpm_filt_Female<-logcpm_filt_Female[-which(duplicated(logcpm_filt_Female)),]
LS_Counts_filt_Female<-LS_Counts_filt_Female[-which(duplicated(LS_Counts_filt_Female)),]

#Dimensions after removing duplicates
dim(LS_Counts_filt_Male)
dim(annot_Male)
dim(logcpm_filt_Male)

dim(LS_Counts_filt_Female)
dim(annot_Female)
dim(logcpm_filt_Female)

sum(rownames(LS_Counts_filt_Male)==annot_Male)
sum(rownames(LS_Counts_filt_Female)==annot_Female)

```

### TMM Normalization
Perform TMM normalization on filtered counts
```{r TMM Norm}
#Make DGEList
Y_Male<- DGEList(counts=LS_Counts_filt_Male, genes = annot_Male$SYMBOL)

Y_norm_Male<-calcNormFactors(Y_Male,method="TMM") #This is the Default Method

Y_Female<- DGEList(counts=LS_Counts_filt_Female, genes = annot_Female$SYMBOL)

Y_norm_Female<-calcNormFactors(Y_Female,method="TMM") #This is the Default Method

#Plot before and after normalization
boxplot(cpm(Y_Male,log=T)[,1:10], main="Male-Before Normalization")

boxplot(cpm(Y_norm_Male,log=T)[,1:10], main="Male-After Normalization")

boxplot(cpm(Y_Female,log=T)[,1:10], main="Female-Before Normalization")

boxplot(cpm(Y_norm_Female,log=T)[,1:10], main="Female-After Normalization")
```
### Fix Exposure Data to include continuous data for those that pass threshold and binary data for those that don't
Also setting the binary OPEs to factors since they had changed to integers
```{r fix exposure data}
OPE_Data_Select_Ready_Male <- OPE_Data_Select_Male[, -1]
rownames(OPE_Data_Select_Ready_Male) <- OPE_Data_Select_Ready_Male$pathways_id
OPE_Data_Select_Ready_Male <- OPE_Data_Select_Ready_Male %>%
  dplyr::select(-pathways_id)
OPE_Data_Select_Ready_Male$dprp_lod<-factor(OPE_Data_Select_Ready_Male$dprp_lod, levels=c("1", "0"))
OPE_Data_Select_Ready_Male$bboep_lod<-factor(OPE_Data_Select_Ready_Male$bboep_lod, levels=c("1", "0"))
OPE_Data_Select_Ready_Male$behp_lod<-factor(OPE_Data_Select_Ready_Male$behp_lod, levels=c("1", "0"))
OPE_Data_Select_Ready_Male$bcpp_lod<-factor(OPE_Data_Select_Ready_Male$bcpp_lod, levels=c("1", "0"))
OPE_Data_Select_Ready_Male$bmpp_lod<-factor(OPE_Data_Select_Ready_Male$bmpp_lod, levels=c("1", "0"))

OPE_Data_Select_Ready_Female <- OPE_Data_Select_Female[, -1]
rownames(OPE_Data_Select_Ready_Female) <- OPE_Data_Select_Ready_Female$pathways_id
OPE_Data_Select_Ready_Female <- OPE_Data_Select_Ready_Female %>%
  dplyr::select(-pathways_id)
OPE_Data_Select_Ready_Female$dprp_lod<-factor(OPE_Data_Select_Ready_Female$dprp_lod, levels=c("1", "0"))
OPE_Data_Select_Ready_Female$bboep_lod<-factor(OPE_Data_Select_Ready_Female$bboep_lod, levels=c("1", "0"))
OPE_Data_Select_Ready_Female$behp_lod<-factor(OPE_Data_Select_Ready_Female$behp_lod, levels=c("1", "0"))
OPE_Data_Select_Ready_Female$bcpp_lod<-factor(OPE_Data_Select_Ready_Female$bcpp_lod, levels=c("1", "0"))
OPE_Data_Select_Ready_Female$bmpp_lod<-factor(OPE_Data_Select_Ready_Female$bmpp_lod, levels=c("1", "0"))

```

### Save filtered/normalized data
```{r save}
#save(Y_norm_Male, OPE_Data_Select_Ready_Male, Covariates_Male, annot_Male, file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/FilteredNormalizedData_Male_102323.RData")

#save(Y_norm_Female, OPE_Data_Select_Ready_Female, Covariates_Female, annot_Female, file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/FilteredNormalizedData_Female_102323.RData")
```

### Quality Control: Principle Components Analysis - Male
Performing a PCA and then testing association of covariates with PCs
```{r Principal Components}
logcpm_Final_Male <- cpm(LS_Counts_filt_Male, log=TRUE)
PCobj_Male =prcomp(t(logcpm_Final_Male), scale=TRUE)
PCs_Male = PCobj_Male$x
PCs_Male =as.data.frame(PCs_Male[,1:50])

rownames(Covariates_Male) <- Covariates_Male$pathways_id
Covariates_PCs_Male<-merge(Covariates_Male, PCs_Male,by='row.names')
rownames(Covariates_PCs_Male)<-Covariates_PCs_Male[,1]
Covariates_PCs_Male<-Covariates_PCs_Male[,-1]

EVAL_Male<-as.data.frame(matrix(NA,nrow=4,ncol=13))
colnames(EVAL_Male)<-c("P_SG", "P_CollectionGestage", "P_MaternalAge", "P_MaternalBMI", "P_MaternalRace", "P_MaternalEducation", "P_Batch", "P_LaborType", "P_DeliveryMethod", "P_Alcohol", "P_Smoking", "P_Parity", "P_Gravidity")

for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~opfr_sg,data=Covariates_PCs_Male)
  EVAL_Male[i,1] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~OPFR_Gestage,data=Covariates_PCs_Male)
  EVAL_Male[i,2] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~h_m_delivery_age,data=Covariates_PCs_Male)
  EVAL_Male[i,3] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}
for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~h_m_prepreg_bmi,data=Covariates_PCs_Male)
  EVAL_Male[i,4] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}
for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~h_m_race,data=Covariates_PCs_Male)
  EVAL_Male[i,5] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~h_m_enroll_educ,data=Covariates_PCs_Male)
  EVAL_Male[i,6] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~batches,data=Covariates_PCs_Male)
  EVAL_Male[i,7] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~cg_labor_type,data=Covariates_PCs_Male)
  EVAL_Male[i,8] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~h_del_method,data=Covariates_PCs_Male)
  EVAL_Male[i,9] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}


for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~h_m_alcohol_prenatal,data=Covariates_PCs_Male)
  EVAL_Male[i,10] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~Smoking_Composite,data=Covariates_PCs_Male)
  EVAL_Male[i,11] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~h_firstborn,data=Covariates_PCs_Male)
  EVAL_Male[i,12] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Male[,i]~h_gravidity,data=Covariates_PCs_Male)
  EVAL_Male[i,13] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

EVAL_Male$PropVar = summary(PCobj_Male)$importance["Proportion of Variance", 1:50]

#Coloring by Variables associated with PC1 or PC2
#PC2 v PC1- Colored by Smoking
plot(PCs_Male$PC1,PCs_Male$PC2,col=as.factor(Covariates_PCs_Male$Smoking_Composite),xlab="PC1",ylab="PC2", main="Maternal Smoking")
#PC2 v PC1- Colored by Maternal Education
plot(PCs_Male$PC1,PCs_Male$PC2,col=Covariates_PCs_Male$h_m_enroll_educ,xlab="PC1",ylab="PC2", main="Maternal Education")
#PC2 v PC1- Colored by Sequencing Batch
plot(PCs_Male$PC1,PCs_Male$PC2,col=as.factor(Covariates_PCs_Male$batches),xlab="PC1",ylab="PC2", main="Sequencing Batches")

#Scree Plot
EVAL_Male$PC <- as.numeric(rownames(EVAL_Male))

EVAL_Male %>%
  ggplot(aes(x=PC, y=PropVar, group=1)) +
  geom_line()+
  geom_point() +
  ggtitle("Variance by PC") + 
  ylab("Proportion of Variance") +
  scale_x_continuous(breaks=seq(0,50, by=1))+
  ylim(0,0.5)
```

### Quality Control: Principle Components Analysis - Female
Performing a PCA and then testing association of covariates with PCs
```{r Principal Components Female}
logcpm_Final_Female <- cpm(LS_Counts_filt_Female, log=TRUE)
PCobj_Female =prcomp(t(logcpm_Final_Female), scale=TRUE)
PCs_Female = PCobj_Female$x
PCs_Female =as.data.frame(PCs_Female[,1:50])

rownames(Covariates_Female) <- Covariates_Female$pathways_id
Covariates_PCs_Female<-merge(Covariates_Female, PCs_Female,by='row.names')
rownames(Covariates_PCs_Female)<-Covariates_PCs_Female[,1]
Covariates_PCs_Female<-Covariates_PCs_Female[,-1]

EVAL_Female<-as.data.frame(matrix(NA,nrow=4,ncol=13))
colnames(EVAL_Female)<-c("P_SG", "P_CollectionGestage", "P_MaternalAge", "P_MaternalBMI", "P_MaternalRace", "P_MaternalEducation", "P_SBatch", "P_LaborType", "P_DeliveryMethod", "P_Alcohol", "P_Smoking", "P_Parity", "P_Gravidity")

for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~opfr_sg,data=Covariates_PCs_Female)
  EVAL_Female[i,1] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~OPFR_Gestage,data=Covariates_PCs_Female)
  EVAL_Female[i,2] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~h_m_delivery_age,data=Covariates_PCs_Female)
  EVAL_Female[i,3] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}
for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~h_m_prepreg_bmi,data=Covariates_PCs_Female)
  EVAL_Female[i,4] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}
for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~h_m_race,data=Covariates_PCs_Female)
  EVAL_Female[i,5] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~h_m_enroll_educ,data=Covariates_PCs_Female)
  EVAL_Female[i,6] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~batches,data=Covariates_PCs_Female)
  EVAL_Female[i,7] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~cg_labor_type,data=Covariates_PCs_Female)
  EVAL_Female[i,8] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~h_del_method,data=Covariates_PCs_Female)
  EVAL_Female[i,9] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}


for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~h_m_alcohol_prenatal,data=Covariates_PCs_Female)
  EVAL_Female[i,10] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~Smoking_Composite,data=Covariates_PCs_Female)
  EVAL_Female[i,11] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~h_firstborn,data=Covariates_PCs_Female)
  EVAL_Female[i,12] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

for (i in 1:50){
  AOV<-aov(PCs_Female[,i]~h_gravidity,data=Covariates_PCs_Female)
  EVAL_Female[i,13] <-summary(AOV)[[1]][["Pr(>F)"]][1]
}

EVAL_Female$PropVar = summary(PCobj_Female)$importance["Proportion of Variance", 1:50]

#Coloring by Variables associated with PC1 or PC2
#PC2 v PC1- Colored by Maternal Age
plot(PCs_Female$PC1,PCs_Female$PC2,col=Covariates_PCs_Female$h_m_delivery_age,xlab="PC1",ylab="PC2", main="Maternal Age")
#PC2 v PC1- Colored by Labor Type
plot(PCs_Female$PC1,PCs_Female$PC2,col=as.factor(Covariates_PCs_Female$cg_labor_type),xlab="PC1",ylab="PC2", main="Labor Type")
#PC2 v PC1- Colored by Labor Type
plot(PCs_Female$PC1,PCs_Female$PC2,col=as.factor(Covariates_PCs_Female$batches),xlab="PC1",ylab="PC2", main="Batch")

#Scree Plot
EVAL_Female$PC <- as.numeric(rownames(EVAL_Female))

EVAL_Female %>%
  ggplot(aes(x=PC, y=PropVar, group=1)) +
  geom_line()+
  geom_point() +
  ggtitle("Variance by PC") + 
  ylab("Proportion of Variance") +
  scale_x_continuous(breaks=seq(0,50, by=1))+
  ylim(0,0.5)
```










