---
title: "WGCNA_SummaryFigure_091824"
author: "Samantha Lapehn"
date: "2024-09-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## WGCNA Summary Figure

This code will be used to create a heatmap summary figure

The results uploaded here are from the CANDLE WGCNA modules that have been associated with OPEs. Sept 2024 Update- Remaking Figure after rerunning WGCNA linear models. 

### Load Packages
```{r load packages}
library(tidyverse)
library(pheatmap)
library(RColorBrewer)
```

### Load Data
Loading .csv files which contain summary results of the models that were significant. 
```{r load data}
ContinuousOPE <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/WGCNA/WGCNASummaryforFigure_ContinuousOPE_091824.csv")
BinaryOPE <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/WGCNA/WGCNASummaryforFigure_BinaryOPE_091824.csv")
```

### Prepare Heatmap input
```{r heatmap annotations}

rownames(ContinuousOPE) <- ContinuousOPE$X

ContinuousOPE <- ContinuousOPE %>%
  dplyr::select(-X)

ContinuousOPE_forheatmap <- t(ContinuousOPE)

rownames(BinaryOPE) <- BinaryOPE$X

BinaryOPE <- BinaryOPE %>%
  dplyr::select(-X)

BinaryOPE_forheatmap <- t(BinaryOPE)

colors <- rev(brewer.pal(11, "RdBu"))


ContinuousOPEannot <- c("BCETP", "BDCPP", "BDCPP", "DPHP", "DBUP & DIBP", "DEP", "Mixture", "Mixture")
ContinuousModel <- c("Male","All", "Male", "All", "Male", "Female", "All", "Male")

Annot_Continuous <- data.frame("OPE"=ContinuousOPEannot, "Sex"=ContinuousModel)
rownames(Annot_Continuous)<-rownames(ContinuousOPE)

BinaryOPEannot <- c("BBOEP", "BBOEP", "BBOEP", "BEHP", "BEHP", "BEHP", "BCPP", "BCPP", "BMPP", "BMPP")
BinaryModel <- c("All", "Male", "Female", "All", "Male", "Female", "All", "Female", "All", "Female")

Annot_Continuous <- data.frame("OPE"=ContinuousOPEannot, "Sex"=ContinuousModel)
rownames(Annot_Continuous)<-rownames(ContinuousOPE)

Annot_Binary <- data.frame("OPE"=BinaryOPEannot, "Sex"=BinaryModel)
rownames(Annot_Binary)<-rownames(BinaryOPE)

SubColors<-brewer.pal(8, "Dark2")
annotation_colors = list(
  OPE = c("BCETP"=SubColors[1],"BDCPP"=SubColors[2],"DBUP & DIBP"=SubColors[3],"DPHP"=SubColors[4], "DEP"=SubColors[6],"Mixture"=SubColors[5]),
  Sex=c(All="Black",Female="deeppink",Male="Blue")
  )

annotation_colors_binary = list(
  OPE = c("BBOEP"=SubColors[1],"BEHP"=SubColors[2],"BCPP"=SubColors[3],"BMPP"=SubColors[4]),
  Sex=c(All="Black",Female="deeppink",Male="Blue")
  )

```

### Continuous OPE Heatmap
```{r Continuous OPE}

Breaks_Cont<-seq(-0.01,0.01, length.out=12)
pheatmap(ContinuousOPE_forheatmap,
         color = colors,
        breaks=Breaks_Cont,
         scale='none',# unscaled data
         na.color="grey",
         cluster_rows =F,
         cluster_cols=F,
         na_col="grey",
         cellwidth=10, #Change tehe for size
         cellheight=20, #change these for siz
         fontsize_col =0.1, 
         fontsize_row=12, 
         annotation_col=Annot_Continuous, 
         annotation_colors = annotation_colors)
```

### Binary OPE Heatmap
```{r Binary OPE}
Breaks_Binary<-seq(-0.015,0.015, length.out=12)
pheatmap(BinaryOPE_forheatmap,
         color = colors,
        breaks=Breaks_Binary,
         scale='none',# unscaled data
         na.color="grey",
         cluster_rows =F,
         cluster_cols=F,
         na_col="grey",
         cellwidth=10, #Change tehe for size
         cellheight=20, #change these for siz
         fontsize_col =0.1, 
         fontsize_row=12, 
         annotation_col=Annot_Binary, 
         annotation_colors = annotation_colors_binary)

#try horizontal figure

pheatmap(BinaryOPE,
         color = colors,
        breaks=Breaks_Binary,
         scale='none',# unscaled data
         na.color="grey",
         cluster_rows =F,
         cluster_cols=F,
         na_col="grey",
         cellwidth=20, #Change tehe for size
         cellheight=10, #change these for siz
         fontsize_row =0.1, 
         fontsize_col=12, 
         annotation_row=Annot_Binary, 
         annotation_colors = annotation_colors_binary)
```

### Upload WGCNA Pathway Results
These are the CANDLE WGCNA enriched pathways identified via Kegga 
```{r upload pathways}
WGCNA_Pathways <- read.csv("/active/paquette_a/CANDLE_WGCNA/module_pathways/allModules_enrichedPathways.csv")

#Filter to only include modules from continuous OPE modules
ContinuousOPE_Modules <- colnames(ContinuousOPE)
ContinuousOPE_Modules <- tolower(ContinuousOPE_Modules)

WGCNA_Pathways_Continuous <- WGCNA_Pathways %>% 
  dplyr::filter(module %in% ContinuousOPE_Modules) %>%
  dplyr::arrange(module)

WGCNA_Pathways_Continuous$module <- as.factor(WGCNA_Pathways_Continuous$module)

ContinuousOrder <- c(rep(1, 2), rep(2, 1), rep(3,9), rep(4,4), rep(5,1))

WGCNA_Pathways_Continuous$ContinuousOrder <- ContinuousOrder

#filter to only include modules from binary OPE modules
BinaryOPE_Modules <- colnames(BinaryOPE)
BinaryOPE_Modules <- tolower(BinaryOPE_Modules)

WGCNA_Pathways_Binary <- WGCNA_Pathways %>% 
  dplyr::filter(module %in% BinaryOPE_Modules) %>%
  dplyr::arrange(module)

WGCNA_Pathways_Binary$module <- as.factor(WGCNA_Pathways_Binary$module)
summary(WGCNA_Pathways_Binary$module)

BinaryOrder <- c(rep(1, 2), rep(2, 1), rep(3,4), rep(4,4), rep(5,4), rep(6,2), rep(7,2), rep(8, 2), rep(9,1), rep(10,10), rep(11,2), rep(12,9), rep(13,2), rep(14,20), rep(15,1))

WGCNA_Pathways_Binary$BinaryOrder <- BinaryOrder

```


### Make WGCNA Pathway Figure
```{r dot plot}
ggplot(WGCNA_Pathways_Continuous, aes(x=module, y=reorder(Pathway.Name, ContinuousOrder))) + 
  geom_point(aes(size = -log10(P.DE), color=Super.Group)) +
  theme_bw() + 
  xlab("WGCNA Module") + 
  ylab("KEGG Pathways")


ggplot(WGCNA_Pathways_Binary, aes(x=module, y=reorder(Pathway.Name, BinaryOrder))) + 
  geom_point(aes(size = -log10(P.DE), color=Super.Group)) +
  theme_bw() + 
  xlab("WGCNA Module") + 
  ylab("KEGG Pathways")
```

