---
title: "WGCNA Indiidual OPEs Linear Models"
author: "Samantha Lapehn"
date: "2023-11-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## WGCNA Individual OPEs Linear Models
WGCNA modules were generated  on the full CANDLE transcriptome dataset (N=794) (Parenti et al. 2024 Journal of Nutrition).

## Load Packages
```{r load packages}
library(tidyverse)
```

## Load  Data
Loading the unmerged MEs here
```{r load data}
load(file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CANDLE_WGCNA_copy_110123/CANDLE_wide_WGCNA_symbol.RData") #Loading version with HGNC symbols

load(file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CANDLE_WGCNA_copy_110123/CANDLE_wide_WGCNA_ensembl.RData") #loading version with ensembl IDs

load(file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/FilteredNormalizedData_102023.RData")
#Note: Loading this data only to pull the OPE values, not to use the expression data

Covariates <- read.csv("/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CleanCovariates_102023.csv")

load(file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CANDLE_WGCNA_copy_110123/CANDLE_wide_WGCNA_ensembl_fullObject.RData") #Loading full ME object just in case

load(file="/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/CANDLE_WGCNA_copy_110123/cqn_data/cqn_data.RData") #Loading CQN Expr Data

```

## Filter MEs to only include participants with OPE Data
```{r filter MEs}
MEs_forOPE <- MEs %>%
  dplyr::filter(row.names(MEs) %in% row.names(OPE_Data_Select_Ready))
```

## Update sample orders to match
```{r check sample order}
reorder <- match(rownames(OPE_Data_Select_Ready), rownames(MEs_forOPE))
ME_reorder <- MEs_forOPE[reorder,]

check <- rownames(ME_reorder) == rownames(OPE_Data_Select_Ready)
sum(check)

check2 <- rownames(ME_reorder) == Covariates$pathways_id
sum(check2)

```

### Fix Data Class
```{r Covar Data Class Fix}
Covariates$h_m_race <- as.factor(Covariates$h_m_race)
Covariates$h_m_enroll_educ <- as.factor(Covariates$h_m_enroll_educ)
Covariates$batches <- as.factor(Covariates$batches)
Covariates$cg_labor_type <- as.factor(Covariates$cg_labor_type)
Covariates$h_del_method <- as.factor(Covariates$h_del_method)
Covariates$h_m_alcohol_prenatal <- as.factor(Covariates$h_m_alcohol_prenatal)
Covariates$Smoking_Composite <- as.factor(Covariates$Smoking_Composite)
Covariates$h_firstborn <- as.factor(Covariates$h_firstborn)
Covariates$h_c_sex <- as.factor(Covariates$h_c_sex)
```


# Running a separate code chunk for each OPE
5 OPEs are log transformed continuous values
5 OPEs are binary (present or absent) due to high percent below LOD

For each OPE, I am running an unadjusted and adjusted linear model for associations with 40 module eigengenes identified through WGCNA. 

## Log DEP
```{r Log DEP}
UnadjustedLM_Results_DEP<-list()
Unadjusted_LM_Pval_DEP <- list()
UnadjustedLM_DEP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_unadjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$log_dep) 
  
  UnadjustedLM_Results_DEP[[i]]<-summary(LM_unadjusted)
  names(UnadjustedLM_Results_DEP)[i]<-colnames(ME_reorder)[i] 
  
  Unadjusted_LM_Pval_DEP[[i]] <- summary(LM_unadjusted)$coefficients[2,4]
  names(Unadjusted_LM_Pval_DEP)[i]<-colnames(ME_reorder)[i]
  
  UnadjustedLM_DEP_coefficients[[i]] <- broom::tidy(LM_unadjusted)
  names(UnadjustedLM_DEP_coefficients)[i]<-colnames(ME_reorder)[i] 
}
#Printing unadjusted p-values
print(Unadjusted_LM_Pval_DEP)

AdjustedLM_Results_DEP<-list()
Adjusted_LM_Pval_DEP <- list()
Adjusted_LM_EE_DEP <- list()
AdjustedLM_DEP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_adjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$log_dep 
                  +Covariates$h_c_sex
                  +Covariates$opfr_sg
                   +Covariates$h_m_delivery_age
                   +Covariates$h_m_prepreg_bmi
                   +Covariates$h_m_race
                   +Covariates$h_m_enroll_educ
                   +Covariates$batches
                   +Covariates$cg_labor_type
                   +Covariates$h_del_method
                   +Covariates$h_m_alcohol_prenatal
                   +Covariates$Smoking_Composite
                   +Covariates$h_firstborn
                   +Covariates$h_gravidity)
  
  AdjustedLM_Results_DEP[[i]]<-summary(LM_adjusted)
  names(AdjustedLM_Results_DEP)[i]<-colnames(ME_reorder)[i] 
  
  AdjustedLM_DEP_coefficients[[i]] <- broom::tidy(LM_adjusted)
  names(AdjustedLM_DEP_coefficients)[i]<-colnames(ME_reorder)[i] 
  
  Adjusted_LM_Pval_DEP[[i]] <- summary(LM_adjusted)$coefficients[2,4]
  names(Adjusted_LM_Pval_DEP)[i]<-colnames(ME_reorder)[i]
  
  Adjusted_LM_EE_DEP[[i]] <- summary(LM_adjusted)$coefficients[2,1]
  names(Adjusted_LM_EE_DEP)[i]<-colnames(ME_reorder)[i]
}

print(Adjusted_LM_Pval_DEP)
```

## Log DBUP_DIBP
```{r DBUP_DIBP}
UnadjustedLM_Results_DBUP_DIBP<-list()
Unadjusted_LM_Pval_DBUP_DIBP <- list()
UnadjustedLM_DBUP_DIBP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_unadjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$log_dbup_dibp) 
  
  UnadjustedLM_Results_DBUP_DIBP[[i]]<-summary(LM_unadjusted)
  names(UnadjustedLM_Results_DBUP_DIBP)[i]<-colnames(ME_reorder)[i] 
  
  Unadjusted_LM_Pval_DBUP_DIBP[[i]] <- summary(LM_unadjusted)$coefficients[2,4]
  names(Unadjusted_LM_Pval_DBUP_DIBP)[i]<-colnames(ME_reorder)[i]
  
  UnadjustedLM_DBUP_DIBP_coefficients[[i]] <- broom::tidy(LM_unadjusted)
  names(UnadjustedLM_DBUP_DIBP_coefficients)[i]<-colnames(ME_reorder)[i] 
}
#Printing unadjusted model p-values
print(Unadjusted_LM_Pval_DBUP_DIBP)

AdjustedLM_Results_DBUP_DIBP<-list()
Adjusted_LM_Pval_DBUP_DIBP <- list()
Adjusted_LM_EE_DBUP_DIBP <- list()
AdjustedLM_DBUP_DIBP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_adjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$log_dbup_dibp 
                  +Covariates$h_c_sex
                   +Covariates$opfr_sg
                   +Covariates$h_m_delivery_age
                   +Covariates$h_m_prepreg_bmi
                   +Covariates$h_m_race
                   +Covariates$h_m_enroll_educ
                   +Covariates$batches
                   +Covariates$cg_labor_type
                   +Covariates$h_del_method
                   +Covariates$h_m_alcohol_prenatal
                   +Covariates$Smoking_Composite
                   +Covariates$h_firstborn
                   +Covariates$h_gravidity)
  
  AdjustedLM_Results_DBUP_DIBP[[i]]<-summary(LM_adjusted)
  names(AdjustedLM_Results_DBUP_DIBP)[i]<-colnames(ME_reorder)[i] 
  
  AdjustedLM_DBUP_DIBP_coefficients[[i]] <- broom::tidy(LM_adjusted)
  names(AdjustedLM_DBUP_DIBP_coefficients)[i]<-colnames(ME_reorder)[i] 
  
  Adjusted_LM_Pval_DBUP_DIBP[[i]] <- summary(LM_adjusted)$coefficients[2,4]
  names(Adjusted_LM_Pval_DBUP_DIBP)[i]<-colnames(ME_reorder)[i]
  
  Adjusted_LM_EE_DBUP_DIBP[[i]] <- summary(LM_adjusted)$coefficients[2,1]
  names(Adjusted_LM_EE_DBUP_DIBP)[i]<-colnames(ME_reorder)[i]
}
#Printing adjusted model p-values
print(Adjusted_LM_Pval_DBUP_DIBP)

```

## Log BCETP
```{r Log BCETP}
UnadjustedLM_Results_BCETP<-list()
Unadjusted_LM_Pval_BCETP <- list()
UnadjustedLM_BCETP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_unadjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$log_bcetp) 
  
  UnadjustedLM_Results_BCETP[[i]]<-summary(LM_unadjusted)
  names(UnadjustedLM_Results_BCETP)[i]<-colnames(ME_reorder)[i] 
  
  Unadjusted_LM_Pval_BCETP[[i]] <- summary(LM_unadjusted)$coefficients[2,4]
  names(Unadjusted_LM_Pval_BCETP)[i]<-colnames(ME_reorder)[i]
  
  UnadjustedLM_BCETP_coefficients[[i]] <- broom::tidy(LM_unadjusted)
  names(UnadjustedLM_BCETP_coefficients)[i]<-colnames(ME_reorder)[i] 
}
#Printing unadjusted model p-values
print(Unadjusted_LM_Pval_BCETP)

AdjustedLM_Results_BCETP<-list()
Adjusted_LM_Pval_BCETP <- list()
Adjusted_LM_EE_BCETP <- list()
AdjustedLM_BCETP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_adjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$log_bcetp
                  +Covariates$h_c_sex
                   +Covariates$opfr_sg
                   +Covariates$h_m_delivery_age
                   +Covariates$h_m_prepreg_bmi
                   +Covariates$h_m_race
                   +Covariates$h_m_enroll_educ
                   +Covariates$batches
                   +Covariates$cg_labor_type
                   +Covariates$h_del_method
                   +Covariates$h_m_alcohol_prenatal
                   +Covariates$Smoking_Composite
                   +Covariates$h_firstborn
                   +Covariates$h_gravidity)
  
  AdjustedLM_Results_BCETP[[i]]<-summary(LM_adjusted)
  names(AdjustedLM_Results_BCETP)[i]<-colnames(ME_reorder)[i] 
  
  AdjustedLM_BCETP_coefficients[[i]] <- broom::tidy(LM_adjusted)
  names(AdjustedLM_BCETP_coefficients)[i]<-colnames(ME_reorder)[i] 
  
  Adjusted_LM_Pval_BCETP[[i]] <- summary(LM_adjusted)$coefficients[2,4]
  names(Adjusted_LM_Pval_BCETP)[i]<-colnames(ME_reorder)[i]
  
  Adjusted_LM_EE_BCETP[[i]] <- summary(LM_adjusted)$coefficients[2,1]
  names(Adjusted_LM_EE_BCETP)[i]<-colnames(ME_reorder)[i]
}
#Printing adjusted model p-values
print(Adjusted_LM_Pval_BCETP)
```

## Log BDCPP
```{r BDCPP}
UnadjustedLM_Results_BDCPP<-list()
Unadjusted_LM_Pval_BDCPP <- list()
UnadjustedLM_BDCPP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_unadjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$log_bdcpp) 
  
  UnadjustedLM_Results_BDCPP[[i]]<-summary(LM_unadjusted)
  names(UnadjustedLM_Results_BDCPP)[i]<-colnames(ME_reorder)[i] 
  
  Unadjusted_LM_Pval_BDCPP[[i]] <- summary(LM_unadjusted)$coefficients[2,4]
  names(Unadjusted_LM_Pval_BDCPP)[i]<-colnames(ME_reorder)[i]
  
  UnadjustedLM_BDCPP_coefficients[[i]] <- broom::tidy(LM_unadjusted)
  names(UnadjustedLM_BDCPP_coefficients)[i]<-colnames(ME_reorder)[i] 
}
#Printing unadjusted model p-values
print(Unadjusted_LM_Pval_BDCPP)

AdjustedLM_Results_BDCPP<-list()
Adjusted_LM_Pval_BDCPP <- list()
Adjusted_LM_EE_BDCPP <- list()
AdjustedLM_BDCPP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_adjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$log_bdcpp
                  +Covariates$h_c_sex
                   +Covariates$opfr_sg
                   +Covariates$h_m_delivery_age
                   +Covariates$h_m_prepreg_bmi
                   +Covariates$h_m_race
                   +Covariates$h_m_enroll_educ
                   +Covariates$batches
                   +Covariates$cg_labor_type
                   +Covariates$h_del_method
                   +Covariates$h_m_alcohol_prenatal
                   +Covariates$Smoking_Composite
                   +Covariates$h_firstborn
                   +Covariates$h_gravidity)
  
  AdjustedLM_Results_BDCPP[[i]]<-summary(LM_adjusted)
  names(AdjustedLM_Results_BDCPP)[i]<-colnames(ME_reorder)[i] 
  
  AdjustedLM_BDCPP_coefficients[[i]] <- broom::tidy(LM_adjusted)
  names(AdjustedLM_BDCPP_coefficients)[i]<-colnames(ME_reorder)[i] 
  
  Adjusted_LM_Pval_BDCPP[[i]] <- summary(LM_adjusted)$coefficients[2,4]
  names(Adjusted_LM_Pval_BDCPP)[i]<-colnames(ME_reorder)[i]
  
  Adjusted_LM_EE_BDCPP[[i]] <- summary(LM_adjusted)$coefficients[2,1]
  names(Adjusted_LM_EE_BDCPP)[i]<-colnames(ME_reorder)[i]
}
#Printing adjusted model p-values
print(Adjusted_LM_Pval_BDCPP)

```
## Log DPHP
```{r Log DPHP}
UnadjustedLM_Results_DPHP<-list()
Unadjusted_LM_Pval_DPHP <- list()
UnadjustedLM_DPHP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_unadjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$log_dphp) 
  
  UnadjustedLM_Results_DPHP[[i]]<-summary(LM_unadjusted)
  names(UnadjustedLM_Results_DPHP)[i]<-colnames(ME_reorder)[i] 
  
  Unadjusted_LM_Pval_DPHP[[i]] <- summary(LM_unadjusted)$coefficients[2,4]
  names(Unadjusted_LM_Pval_DPHP)[i]<-colnames(ME_reorder)[i]
  
  UnadjustedLM_DPHP_coefficients[[i]] <- broom::tidy(LM_unadjusted)
  names(UnadjustedLM_DPHP_coefficients)[i]<-colnames(ME_reorder)[i] 
}
#Printing unadjusted model p-values
print(Unadjusted_LM_Pval_DPHP)

AdjustedLM_Results_DPHP<-list()
Adjusted_LM_Pval_DPHP <- list()
Adjusted_LM_EE_DPHP <- list()
AdjustedLM_DPHP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_adjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$log_dphp
                  +Covariates$h_c_sex
                   +Covariates$opfr_sg
                   +Covariates$h_m_delivery_age
                   +Covariates$h_m_prepreg_bmi
                   +Covariates$h_m_race
                   +Covariates$h_m_enroll_educ
                   +Covariates$batches
                   +Covariates$cg_labor_type
                   +Covariates$h_del_method
                   +Covariates$h_m_alcohol_prenatal
                   +Covariates$Smoking_Composite
                   +Covariates$h_firstborn
                   +Covariates$h_gravidity)
  
  AdjustedLM_Results_DPHP[[i]]<-summary(LM_adjusted)
  names(AdjustedLM_Results_DPHP)[i]<-colnames(ME_reorder)[i] 
  
  AdjustedLM_DPHP_coefficients[[i]] <- broom::tidy(LM_adjusted)
  names(AdjustedLM_DPHP_coefficients)[i]<-colnames(ME_reorder)[i] 
  
  Adjusted_LM_Pval_DPHP[[i]] <- summary(LM_adjusted)$coefficients[2,4]
  names(Adjusted_LM_Pval_DPHP)[i]<-colnames(ME_reorder)[i]
  
  Adjusted_LM_EE_DPHP[[i]] <- summary(LM_adjusted)$coefficients[2,1]
  names(Adjusted_LM_EE_DPHP)[i]<-colnames(ME_reorder)[i]
}
#Printing adjusted model p-values
print(Adjusted_LM_Pval_DPHP)
```

## Binary dprp
```{r Binary dprp}
UnadjustedLM_Results_DPRP<-list()
Unadjusted_LM_Pval_DPRP <- list()
UnadjustedLM_DPRP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_unadjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$dprp_lod) 
  
  UnadjustedLM_Results_DPRP[[i]]<-summary(LM_unadjusted)
  names(UnadjustedLM_Results_DPRP)[i]<-colnames(ME_reorder)[i] 
  
  Unadjusted_LM_Pval_DPRP[[i]] <- summary(LM_unadjusted)$coefficients[2,4]
  names(Unadjusted_LM_Pval_DPRP)[i]<-colnames(ME_reorder)[i]
  
  UnadjustedLM_DPRP_coefficients[[i]] <- broom::tidy(LM_unadjusted)
  names(UnadjustedLM_DPRP_coefficients)[i]<-colnames(ME_reorder)[i] 
}
#Printing unadjusted model p-values
print(Unadjusted_LM_Pval_DPRP)

AdjustedLM_Results_DPRP<-list()
Adjusted_LM_Pval_DPRP <- list()
Adjusted_LM_EE_DPRP <- list()
AdjustedLM_DPRP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_adjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$dprp_lod
                  +Covariates$h_c_sex 
                  +Covariates$opfr_sg
                   +Covariates$h_m_delivery_age
                   +Covariates$h_m_prepreg_bmi
                   +Covariates$h_m_race
                   +Covariates$h_m_enroll_educ
                   +Covariates$batches
                   +Covariates$cg_labor_type
                   +Covariates$h_del_method
                   +Covariates$h_m_alcohol_prenatal
                   +Covariates$Smoking_Composite
                   +Covariates$h_firstborn
                   +Covariates$h_gravidity)
  
  AdjustedLM_Results_DPRP[[i]]<-summary(LM_adjusted)
  names(AdjustedLM_Results_DPRP)[i]<-colnames(ME_reorder)[i] 
  
  AdjustedLM_DPRP_coefficients[[i]] <- broom::tidy(LM_adjusted)
  names(AdjustedLM_DPRP_coefficients)[i]<-colnames(ME_reorder)[i] 
  
  Adjusted_LM_Pval_DPRP[[i]] <- summary(LM_adjusted)$coefficients[2,4]
  names(Adjusted_LM_Pval_DPRP)[i]<-colnames(ME_reorder)[i]
  
  Adjusted_LM_EE_DPRP[[i]] <- summary(LM_adjusted)$coefficients[2,1]
  names(Adjusted_LM_EE_DPRP)[i]<-colnames(ME_reorder)[i]
}
#Printing adjusted model p-values
print(Adjusted_LM_Pval_DPRP)
```

## Binary BBOEP
```{r Binary BBOEP}
UnadjustedLM_Results_BBOEP<-list()
Unadjusted_LM_Pval_BBOEP <- list()
UnadjustedLM_BBOEP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_unadjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$bboep_lod) 
  
  UnadjustedLM_Results_BBOEP[[i]]<-summary(LM_unadjusted)
  names(UnadjustedLM_Results_BBOEP)[i]<-colnames(ME_reorder)[i] 
  
  Unadjusted_LM_Pval_BBOEP[[i]] <- summary(LM_unadjusted)$coefficients[2,4]
  names(Unadjusted_LM_Pval_BBOEP)[i]<-colnames(ME_reorder)[i]
  
  UnadjustedLM_BBOEP_coefficients[[i]] <- broom::tidy(LM_unadjusted)
  names(UnadjustedLM_BBOEP_coefficients)[i]<-colnames(ME_reorder)[i] 
}
#Printing unadjusted model p-values
print(Unadjusted_LM_Pval_BBOEP)

AdjustedLM_Results_BBOEP<-list()
Adjusted_LM_Pval_BBOEP <- list()
Adjusted_LM_EE_BBOEP <- list()
AdjustedLM_BBOEP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_adjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$bboep_lod
                  +Covariates$h_c_sex
                   +Covariates$opfr_sg
                   +Covariates$h_m_delivery_age
                   +Covariates$h_m_prepreg_bmi
                   +Covariates$h_m_race
                   +Covariates$h_m_enroll_educ
                   +Covariates$batches
                   +Covariates$cg_labor_type
                   +Covariates$h_del_method
                   +Covariates$h_m_alcohol_prenatal
                   +Covariates$Smoking_Composite
                   +Covariates$h_firstborn
                   +Covariates$h_gravidity)
  
  AdjustedLM_Results_BBOEP[[i]]<-summary(LM_adjusted)
  names(AdjustedLM_Results_BBOEP)[i]<-colnames(ME_reorder)[i] 
  
  AdjustedLM_BBOEP_coefficients[[i]] <- broom::tidy(LM_adjusted)
  names(AdjustedLM_BBOEP_coefficients)[i]<-colnames(ME_reorder)[i] 
  
  Adjusted_LM_Pval_BBOEP[[i]] <- summary(LM_adjusted)$coefficients[2,4]
  names(Adjusted_LM_Pval_BBOEP)[i]<-colnames(ME_reorder)[i]
  
  Adjusted_LM_EE_BBOEP[[i]] <- summary(LM_adjusted)$coefficients[2,1]
  names(Adjusted_LM_EE_BBOEP)[i]<-colnames(ME_reorder)[i]
}
#Printing adjusted model p-values
print(Adjusted_LM_Pval_BBOEP)
```

## Binary BEHP
```{r Binary BEHP}
UnadjustedLM_Results_BEHP<-list()
Unadjusted_LM_Pval_BEHP <- list()
UnadjustedLM_BEHP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_unadjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$behp_lod) 
  
  UnadjustedLM_Results_BEHP[[i]]<-summary(LM_unadjusted)
  names(UnadjustedLM_Results_BEHP)[i]<-colnames(ME_reorder)[i] 
  
  Unadjusted_LM_Pval_BEHP[[i]] <- summary(LM_unadjusted)$coefficients[2,4]
  names(Unadjusted_LM_Pval_BEHP)[i]<-colnames(ME_reorder)[i]
  
  UnadjustedLM_BEHP_coefficients[[i]] <- broom::tidy(LM_unadjusted)
  names(UnadjustedLM_BEHP_coefficients)[i]<-colnames(ME_reorder)[i] 
}
#Printing unadjusted model p-values
print(Unadjusted_LM_Pval_BEHP)

AdjustedLM_Results_BEHP<-list()
Adjusted_LM_Pval_BEHP <- list()
Adjusted_LM_EE_BEHP <- list()
AdjustedLM_BEHP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_adjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$behp_lod
                  +Covariates$h_c_sex
                   +Covariates$opfr_sg
                   +Covariates$h_m_delivery_age
                   +Covariates$h_m_prepreg_bmi
                   +Covariates$h_m_race
                   +Covariates$h_m_enroll_educ
                   +Covariates$batches
                   +Covariates$cg_labor_type
                   +Covariates$h_del_method
                   +Covariates$h_m_alcohol_prenatal
                   +Covariates$Smoking_Composite
                   +Covariates$h_firstborn
                   +Covariates$h_gravidity)
  
  AdjustedLM_Results_BEHP[[i]]<-summary(LM_adjusted)
  names(AdjustedLM_Results_BEHP)[i]<-colnames(ME_reorder)[i] 
  
  AdjustedLM_BEHP_coefficients[[i]] <- broom::tidy(LM_adjusted)
  names(AdjustedLM_BEHP_coefficients)[i]<-colnames(ME_reorder)[i] 
  
  Adjusted_LM_Pval_BEHP[[i]] <- summary(LM_adjusted)$coefficients[2,4]
  names(Adjusted_LM_Pval_BEHP)[i]<-colnames(ME_reorder)[i]
  
  Adjusted_LM_EE_BEHP[[i]] <- summary(LM_adjusted)$coefficients[2,1]
  names(Adjusted_LM_EE_BEHP)[i]<-colnames(ME_reorder)[i]
}
#Printing adjusted model p-values
print(Adjusted_LM_Pval_BEHP)
```

## Binary BCPP
```{r Binary BCPP}
UnadjustedLM_Results_BCPP<-list()
Unadjusted_LM_Pval_BCPP <- list()
UnadjustedLM_BCPP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_unadjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$bcpp_lod) 
  
  UnadjustedLM_Results_BCPP[[i]]<-summary(LM_unadjusted)
  names(UnadjustedLM_Results_BCPP)[i]<-colnames(ME_reorder)[i] 
  
  Unadjusted_LM_Pval_BCPP[[i]] <- summary(LM_unadjusted)$coefficients[2,4]
  names(Unadjusted_LM_Pval_BCPP)[i]<-colnames(ME_reorder)[i]
  
  UnadjustedLM_BCPP_coefficients[[i]] <- broom::tidy(LM_unadjusted)
  names(UnadjustedLM_BCPP_coefficients)[i]<-colnames(ME_reorder)[i] 
}
#Printing unadjusted model p-values
print(Unadjusted_LM_Pval_BCPP)

AdjustedLM_Results_BCPP<-list()
Adjusted_LM_Pval_BCPP <- list()
Adjusted_LM_EE_BCPP <- list()
AdjustedLM_BCPP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_adjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$bcpp_lod
                  +Covariates$h_c_sex
                   +Covariates$opfr_sg
                   +Covariates$h_m_delivery_age
                   +Covariates$h_m_prepreg_bmi
                   +Covariates$h_m_race
                   +Covariates$h_m_enroll_educ
                   +Covariates$batches
                   +Covariates$cg_labor_type
                   +Covariates$h_del_method
                   +Covariates$h_m_alcohol_prenatal
                   +Covariates$Smoking_Composite
                   +Covariates$h_firstborn
                   +Covariates$h_gravidity)
  
  AdjustedLM_Results_BCPP[[i]]<-summary(LM_adjusted)
  names(AdjustedLM_Results_BCPP)[i]<-colnames(ME_reorder)[i] 
  
  AdjustedLM_BCPP_coefficients[[i]] <- broom::tidy(LM_adjusted)
  names(AdjustedLM_BCPP_coefficients)[i]<-colnames(ME_reorder)[i] 
  
  Adjusted_LM_Pval_BCPP[[i]] <- summary(LM_adjusted)$coefficients[2,4]
  names(Adjusted_LM_Pval_BCPP)[i]<-colnames(ME_reorder)[i]
  
  Adjusted_LM_EE_BCPP[[i]] <- summary(LM_adjusted)$coefficients[2,1]
  names(Adjusted_LM_EE_BCPP)[i]<-colnames(ME_reorder)[i]
}
#Printing adjusted model p-values
print(Adjusted_LM_Pval_BCPP)

```

## Binary BMPP
```{r binary BMPP}
UnadjustedLM_Results_BMPP<-list()
Unadjusted_LM_Pval_BMPP <- list()
UnadjustedLM_BMPP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_unadjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$bmpp_lod) 
  
  UnadjustedLM_Results_BMPP[[i]]<-summary(LM_unadjusted)
  names(UnadjustedLM_Results_BMPP)[i]<-colnames(ME_reorder)[i] 
  
  Unadjusted_LM_Pval_BMPP[[i]] <- summary(LM_unadjusted)$coefficients[2,4]
  names(Unadjusted_LM_Pval_BMPP)[i]<-colnames(ME_reorder)[i]
  
  UnadjustedLM_BMPP_coefficients[[i]] <- broom::tidy(LM_unadjusted)
  names(UnadjustedLM_BMPP_coefficients)[i]<-colnames(ME_reorder)[i] 
}
#Printing unadjusted model p-values
print(Unadjusted_LM_Pval_BMPP)

AdjustedLM_Results_BMPP<-list()
Adjusted_LM_Pval_BMPP <- list()
Adjusted_LM_EE_BMPP <- list()
AdjustedLM_BMPP_coefficients <- list()
for(i in 1:dim(ME_reorder)[2]){
  LM_adjusted<-lm((ME_reorder[,i])~OPE_Data_Select_Ready$bmpp_lod
                  +Covariates$h_c_sex
                   +Covariates$opfr_sg
                   +Covariates$h_m_delivery_age
                   +Covariates$h_m_prepreg_bmi
                   +Covariates$h_m_race
                   +Covariates$h_m_enroll_educ
                   +Covariates$batches
                   +Covariates$cg_labor_type
                   +Covariates$h_del_method
                   +Covariates$h_m_alcohol_prenatal
                   +Covariates$Smoking_Composite
                   +Covariates$h_firstborn
                   +Covariates$h_gravidity)
  
  AdjustedLM_Results_BMPP[[i]]<-summary(LM_adjusted)
  names(AdjustedLM_Results_BMPP)[i]<-colnames(ME_reorder)[i] 
  
  AdjustedLM_BMPP_coefficients[[i]] <- broom::tidy(LM_adjusted)
  names(AdjustedLM_BMPP_coefficients)[i]<-colnames(ME_reorder)[i] 
  
  Adjusted_LM_Pval_BMPP[[i]] <- summary(LM_adjusted)$coefficients[2,4]
  names(Adjusted_LM_Pval_BMPP)[i]<-colnames(ME_reorder)[i]
  
   Adjusted_LM_EE_BMPP[[i]] <- summary(LM_adjusted)$coefficients[2,1]
  names(Adjusted_LM_EE_BMPP)[i]<-colnames(ME_reorder)[i]
}
#Printing adjusted model p-values
print(Adjusted_LM_Pval_BMPP)

```

## Save Adjusted model p-values into Dataframe
```{r save adjusted model pvals}
DEP_pvals <- unlist(Adjusted_LM_Pval_DEP)
DBUP_DIBP_pvals <- unlist(Adjusted_LM_Pval_DBUP_DIBP)
BCETP_pvals <- unlist(Adjusted_LM_Pval_BCETP)
BDCPP_pvals <- unlist(Adjusted_LM_Pval_BDCPP)
DPHP_pvals <- unlist(Adjusted_LM_Pval_DPHP)
DPRP_pvals <- unlist(Adjusted_LM_Pval_DPRP)
BBOEP_pvals <- unlist(Adjusted_LM_Pval_BBOEP)
BEHP_pvals <- unlist(Adjusted_LM_Pval_BEHP)
BCPP_pvals <- unlist(Adjusted_LM_Pval_BCPP)
BMPP_pvals <- unlist(Adjusted_LM_Pval_BMPP)

AdjustedLM_Pval_Summary <- data.frame(DEP_pvals, DBUP_DIBP_pvals, BCETP_pvals, BDCPP_pvals, DPHP_pvals, DPRP_pvals, BBOEP_pvals, BEHP_pvals, BCPP_pvals, BMPP_pvals)

#Save pval summary
#write.csv(AdjustedLM_Pval_Summary, "/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/WGCNA/AdjustedLM_pval_summary_083024.csv")
```

## Save Adjusted model effect estimates into Dataframe
```{r save adjusted model effect estimates}
DEP_EE <- unlist(Adjusted_LM_EE_DEP)
DBUP_DIBP_EE <- unlist(Adjusted_LM_EE_DBUP_DIBP)
BCETP_EE <- unlist(Adjusted_LM_EE_BCETP)
BDCPP_EE <- unlist(Adjusted_LM_EE_BDCPP)
DPHP_EE <- unlist(Adjusted_LM_EE_DPHP)
DPRP_EE <- unlist(Adjusted_LM_EE_DPRP)
BBOEP_EE <- unlist(Adjusted_LM_EE_BBOEP)
BEHP_EE <- unlist(Adjusted_LM_EE_BEHP)
BCPP_EE <- unlist(Adjusted_LM_EE_BCPP)
BMPP_EE <- unlist(Adjusted_LM_EE_BMPP)

AdjustedLM_EE_Summary <- data.frame(DEP_EE, DBUP_DIBP_EE, BCETP_EE, BDCPP_EE, DPHP_EE, DPRP_EE, BBOEP_EE, BEHP_EE, BCPP_EE, BMPP_EE)

#Save effect estimate summary
#write.csv(AdjustedLM_EE_Summary, "/active/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Results/WGCNA/AdjustedLM_EE_summary_083024.csv")
```



















