---
title: "WGCNA_TFenrichment_091924"
author: "Samantha Lapehn"
date: "2024-09-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## WGCNA TF Enrichment

In this code, I will upload the enriched TFs for the modules of interest from the OPE analysis and evaluate whether any of the TFs are nuclear hormone receptors. Sept 2024 Update- fixing to reflect updated data. 

### Load Packages
```{r load packages}
library(tidyverse)
```

### Load Data

1) TF enrichment for all CANDLE WGCNA modules
2) List of TFs that are NHRs

```{r load data}
CANDLE_WGCNA_TFenrichment <- read.csv("/Volumes/paquette_a/CANDLE_WGCNA/module_TFs/allModules_enrichedTFs.csv")
NHR_TFs <- read.csv("/Volumes/paquette_a/slapehn/OPE_and_PlacentalTranscriptome/Data/NHR_TFs_GuidetoPharmacology.csv")
```

### Select modules of interest for Continuous OPEs
```{r modules of interest-continuous OPEs}
#Vector of modules names that are associated with a continuous OPE or QGComp mixture in a full or sex-stratified model
MOI <- c("darkgrey", "pink", "salmon", "midnightblue", "skyblue", "yellow", "brown", "cyan", "darkred", "greenyellow", "royalblue", "white", "grey60", "sienna3")

#Filtering to only include the enrichment results for the modules of interest defined above
OPE_moduleofinterest <- CANDLE_WGCNA_TFenrichment %>%
  dplyr::filter(enriched_in_module %in% MOI)

#Filtering further to only include TFs that are nuclear hormone receptors enriched for the modules of interest defined above
OPE_moduleofinterest_NHRs <- OPE_moduleofinterest %>%
  dplyr::filter(TF %in% NHR_TFs$NHR) 

```

### Dot Plot of Enriched NHRs for Continuous OPEs
without home module colors
```{r NHR enrichment}
OPE_moduleofinterest_NHRs$enriched_in_module <-as.factor(OPE_moduleofinterest_NHRs$enriched_in_module)
summary(OPE_moduleofinterest_NHRs$enriched_in_module)
#Make vector to order NHRs by module of enrichment
Order <- c(rep(1, 1), rep(2, 1), rep(3,2), rep(4,2), rep(5,3), rep(6, 2), rep(7, 3), rep(8, 3), rep(9, 2), rep(10, 3), rep(11, 2))

ggplot(OPE_moduleofinterest_NHRs, aes(x=enriched_in_module, y=reorder(TF, Order), fill=enriched_in_module)) + 
  geom_point(aes(size = -log10(FisherAdjustP)), shape=21) +
  theme_bw() + 
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 5)) + 
  xlab("Module of Enrichment") + 
  scale_fill_manual(values=c("brown"="tan4", "cyan"="cyan","darkgrey"="darkgrey", "darkred"="red3", "midnightblue"="midnightblue", "greenyellow"="greenyellow", "pink"="pink", "salmon"="salmon", "skyblue"="skyblue", "white"="ivory2", "yellow"="yellow")) +
  ylab("Nuclear Hormone Receptor Transcription Factor")

```

### Select modules of interest for Binary OPEs
```{r modules of interest-binary OPEs}
#Vector of modules names that are associated with a continuous OPE or QGComp mixture in a full or sex-stratified model
MOI_Binary <- c("blue", "green", "greenyellow", "lightgreen", "pink", "salmon", "lightcyan",  "brown", "cyan", "darkorange", "magenta", "plum1", "saddlebrown", "tan", "white", "yellowgreen",  "darkturquoise", "orangered4", "steelblue", "darkmagenta", "midnightblue", "red", "turquoise", "paleturquoise", "darkolivegreen")

#Filtering to only include the enrichment results for the modules of interest defined above
OPE_moduleofinterest_binary <- CANDLE_WGCNA_TFenrichment %>%
  dplyr::filter(enriched_in_module %in% MOI_Binary)

#Filtering further to only include TFs that are nuclear hormone receptors enriched for the modules of interest defined above
OPE_moduleofinterest_NHRs_Binary <- OPE_moduleofinterest_binary %>%
  dplyr::filter(TF %in% NHR_TFs$NHR) 

```


### Dot Plot of Enriched NHRs for Binary OPEs
without home module coloring
```{r NHR enrichment Binary}
OPE_moduleofinterest_NHRs_Binary$enriched_in_module <-as.factor(OPE_moduleofinterest_NHRs_Binary$enriched_in_module)
summary(OPE_moduleofinterest_NHRs_Binary$enriched_in_module)
#Make vector to order NHRs by module of enrichment
Order_Binary <- c(rep(1, 1), rep(2, 1), rep(3,3), rep(4,3), rep(5,1), rep(6, 3), rep(7, 4), rep(8, 2), rep(9, 2), rep(10, 2), rep(11, 1), rep(12, 3), rep(13,2), rep(14,3), rep(15,1), rep(16,1), rep(17,2), rep(18,3), rep(19,3))

ggplot(OPE_moduleofinterest_NHRs_Binary, aes(x=enriched_in_module, y=reorder(TF, Order_Binary), fill=enriched_in_module)) + 
  geom_point(aes(size = -log10(FisherAdjustP)), shape=21) +
  theme_bw() + 
  scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 5)) + 
  xlab("Module of Enrichment") + scale_fill_manual(values=c("brown"="tan4", "cyan"="cyan", "darkorange"="darkorange", "darkturquoise"="darkturquoise", "lightcyan"="lightcyan", "green"="forestgreen", "greenyellow"="greenyellow", "lightgreen"="lightgreen", "magenta"="magenta", "midnightblue"="midnightblue", "orangered4"="orangered4", "pink"="pink", "red"="red", "salmon"="salmon", "steelblue"="steelblue", "tan"="tan", "turquoise"="turquoise", "white"="ivory2",  "yellowgreen"="yellowgreen")) +
  ylab("Nuclear Hormone Receptor Transcription Factor")

```

### Total number of enriched TFs
```{r TF counts}
#Total number of TFs enriched for Modules associated with Continuous OPEs
length(unique(OPE_moduleofinterest$TF))

#Total number of TFs enriched for Modules associated with Binary OPEs
length(unique(OPE_moduleofinterest_binary$TF))

#Total number of NHR TFs enriched for Modules associated with continuous OPEs
length(unique(OPE_moduleofinterest_NHRs$TF))

#Total number of NHR TFs enriched for Modules associated with binary OPEs
length(unique(OPE_moduleofinterest_NHRs_Binary$TF))
```


